---
title: "eds212-day4-comp1"
format: html
editor: visual
---

# Create Vectors
```{r}
rain <-  c(3,5,7,9)
#making a vector and storing it in rain
rain_scaled <- 5 * rain
#this stores the scalar rain vector
```
for git messages, you want them to be concise, meaningful, and searchable

# Create matrices

```{r}
my_values <- seq(from = 1, to = 10)
#make a sequence
```

```{r}
my_matrix <- matrix(data = my_values, nrow = 2, ncol = 5, byrow = TRUE)
my_matrix
#create a matrix by using data. 'byrow = TRUE/FALSE' will show how the numbers are laid out
```

rm() code will take it out of our environment

```{r}
my_matrix2 <- matrix(data = my_values, nrow = 3, ncol = 4, byrow = TRUE)
```
## matrix map
```{r}
matrixA <- matrix(data = seq(from = 2, to = 5, length = 10), 
                  nrow = 2, 
                  ncol =5, 
                  byrow = TRUE)
#making another matrix
```

```{r}
#adding or subtracting matrix
my_matrix + matrixA
```
#Matrix Multiplication
```{r}
cats <- matrix(data = c(1,3,5,7), 
               nrow = 2, 
               ncol = 2, 
               byrow = TRUE)
#create matrix of cats
dogs <- matrix(data = c(2,4,6,8), 
               nrow = 2, 
               ncol = 2, 
               byrow = TRUE)
#create matrix of dogs
cats %*% dogs
```

#Leslie matrix example
```{r}
insect_leslie <- matrix(data = c(0, 0 ,600 ,0.2 ,0 ,0 ,0 ,0.08 ,0), 
                        nrow = 3, 
                        ncol = 3,
                        byrow = TRUE)
#create leslie matrix
insect_y0 <- c(12000, 700, 500)
#setting up our initial condition
```

## Find the population structure for each life stage at t = 1 year:
```{r}
#we are now using the dot product to find the population at year 1
insect_y1 <- insect_leslie %*% insect_y0
insect_y1
```
## Find population structure at year 1
```{r}
insect_y2 <- insect_leslie %*% insect_y1
insect_y2
```

notes to self: if you wanted to input just the year, i think you would have to add a forloop
  update: that is what allison said :D








